TacticToe: Learning the Game of Tactics

### 2018-01-29 tactictoe_jv.tex
contribs

\todoi{add prediction of 3 kinds of things}
\begin{itemize}
\item Proof recording at the tactic level is made more precise. Pattern 
matching constructions, opening of submodules are supported. The different 
recording steps (parsing, unfolding and replaying) are regrouped into a single 
\sml function. So, our recording is more complete and fully automatized.
\item Two types of arguments (list of theorems and terms) can be predicted 
independently of the tactic. Combinatorial explosion in the number of tactics 
is prevented through the introduction of abstracted tactics. As desired, their 
instantiations is dependent on the target goal.
\item MCTS (Monte Carlo tree search) now replaces A* in our proof search 
algorithm. MCTS gives a more balanced feedback to guide proof search by 
comparing subtrees instead of leafs. The policy and evaluation are learned 
through supervised learning.

\item First evaluation of \tactictoe and \eprover success rates relative to 
the length of the original proofs.
\item User interaction is enhanced by the minimization and prettification 
applied on the generated proof.
\item The internal ATP \metis used during proof search can be complemented 
by asynchronous calls to \eprover.
\end{itemize}

@@

In order to be able to describe the different components of \tactictoe, 
we need to give to describe what a search tree is, what are its elements and 
how they are organized relative to one other. The following description is 
static and the evolution of the search tree toward a proof is given in 
Section~\ref{sec:proofsearch}.

\begin{definition}(search tree)\\
A search tree $\mathfrak{T}$ is a sixtuple 
$(\mathbb{T},\mathbb{G},\mathbb{A},T,G,A)$ 
that respects the following conditions:
\begin{itemize}
\item $\mathbb{T}$ is a set of tactics, $\mathbb{G}$ is a set of goals 
 and $\mathbb{A}$ is a set of nodes. These sets contains only objects 
 appearing in the search tree.
\item $T$ is a total function from $\mathbb{G}$ to $P(\mathbb{T})$. It takes a 
goal and return the set of tactics already applied to this goal.
\item $G$ is a total function from $\mathbb{A}$ to $P(\mathbb{G})$. It takes a 
node a return the set of goals in this node.
\item $Support(A) = \lbrace (g,t)\ |\ t \in T(g) \rbrace$.
\item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to 
$\mathbb{A}$. It takes a goal and a tactic and returns the node produced by the 
application of the tactic to the goal, if this tactic has been applied to this 
goal.
\item if $t \in T(g)$ and $A(g,t) = a$ then $t(g) = G(a)$.
\end{itemize}

Additional constraints such as acyclicity and the fact that there is exactly 
one root node are properties of this graph that are preserved during the 
application of the MCTS algorithm. That is why we call it a tree although this 
is not specified in this definition.
\end{definition}

If no explicit order is given, we assume that any set is equipped with an 
arbitrary order. In this way, we can access the nth member of set $s$ by 
writing $s_n$. In Figure~\ref{fig:choice}, $(G(a_j))_i$ is written $g_j^i$.

@@

Starting from the bottom parent node, we can construct children of a goal of 
this node by choosing a goal $g_i^0$ on this node and choosing tactics 
$t_1,t_2,\ldots$ to 
apply to this goal. Here the result produces new nodes each containing a list 
of goals to be proven. In each node of the proof tree, a goal can either be 
solved, open or pending (see Definition~\ref{def:solved}).

In the following, new concepts and properties are defined in the context of a 
search tree $\mathfrak{T}$.

@@

\section{Prediction}\label{s:prediction}
The learning-based selection of relevant lemmas significantly improves the 
automation for hammers~\cite{BlanchetteGKKU16}. In \tactictoe, we adapt 
one of the strongest lemma selection 
methods: the modified distance-weighted \emph{$k$ nearest-neighbour} 
classifier~\cite{ckju-pxtp13,DudaniS76} to predict three different objects: 
 tactics, theorems and list of goals.

@@

We now concentrate on the definition a reasonable evaluation function for the 
output of tactics. Theoretically, an interesting list of goal is one 
that has a short proof, and we could measure the likelihood of being 
provable in a number of steps from previous proof search. However, this 
approach produces a lot of negatives\todo{CK: cite hard negative mining 
or/and deep math}. So to reduce this number, we limit 
ourselves to collect output of tactics tested during orthogonalization. We 
declare a list of goals $l$ to be positive if it is has been produced by the 
winner of an orthogonalization competition. The predictor with radius $r$ 
\todo{define radius} produces a set of $r$ list of goals closest to $l$. The 
number of positive example in this set is noted $Positive(r,l)$.
During the proof search, we evaluate nodes through the list of goals they 
contain using the following evaluation function:
\begin{align*}
PriorEvaluation_r (a_i) &= \frac{Positive(r,G(a_i))}{r}\\
\end{align*}

@@

Such a prover already exists in \holfour and is called \metis. It is already 
recognized by the tactic selection mechanism in \tactictoe and can have its 
premises predicted dynamically when in an abstracted form. Nevertheless,
we think that the performance of \tactictoe can be boosted by giving the ATP 
\metis a special status. This arrangement consists in always predicting 
premises for \metis, giving it a slightly higher timeout and 
always trying it first on each node. These modifications only induce an
overhead linear on the number nodes as a general purpose prover does not create
goal obligations. \todoi{Define goal obligation in the introduction}

>>>>

\begin{itemize}
\item Proof recording at the tactic level is made more precise. Pattern 
matching constructions, opening of submodules are supported. The different 
recording steps (parsing, unfolding and replaying) are regrouped into a single 
\sml function.
\item MCTS (Monte Carlo tree search) now replaces A* in our proof search 
algorithm. MCTS gives a more balanced feedback to guide proof search by 
comparing subtrees instead of leafs. The policy and evaluation are learned 
through supervised learning.
\item Proof guidance required by the MCTS algorithm is given by predictors for 
three kind of objects: tactics, theorems and list of goals.
\item The orthogonalization process that eliminates redundant tactics is 
formally described.
\item A tactic abstraction method is introduced. It enable us to create 
more general and flexible tactics where tactic arguments are dynamically 
predicted.
\item The internal ATP \metis is complemented 
by asynchronous calls to \eprover to help \tactictoe during proof search.
\todoi{More on evaluation??}
\item Evaluation of \tactictoe success rate relative to 
the length of the original proofs.
\item User interaction is enhanced by the minimization and prettification 
applied on the generated proof.
\end{itemize}

@@

The different components of \tactictoe revolves around the concept of a 
search tree. The following definitions are made relative to a static search 
tree, the description of its evolution during proof search will be given in 
Section~\ref{sec:proofsearch}.

\begin{definition}(search tree)\\
A search tree $\mathfrak{T}$ is a sixtuple 
$(\mathbb{T},\mathbb{G},\mathbb{A},T,G,A)$ 
that respects the following conditions:
\begin{itemize}
\item $\mathbb{T}$ is a set of tactics, $\mathbb{G}$ is a set of goals 
 and $\mathbb{A}$ is a set of nodes. These sets contains only objects 
 appearing in the search tree.
\item $T$ is a total function from $\mathbb{G}$ to $P(\mathbb{T})$. It takes a 
goal and return the set of tactics already applied to this goal.
\item $G$ is a total function from $\mathbb{A}$ to $P(\mathbb{G})$. It takes a 
node a return the set of goals in this node.

\item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to 
$\mathbb{A}$. It takes a goal and a tactic and returns the node produced by the 
application of the tactic to the goal, if this tactic has been applied to this 
goal. Its support $Support(A)$ is by definition $\lbrace (g,t)\ |\ t \in T(g) 
\rbrace$.
\item if $(g,t) \in Support (A)$ and $A(g,t) = a$, then $t(g) = G(a)$.
\end{itemize}

Additional constraints such as acyclicity and the fact that there is exactly 
one root node are properties of this graph that are preserved during the 
application of the MCTS algorithm. That is why we call it a tree, although this 
is not specified in this definition.
\end{definition}

If no explicit order is given, we assume that any set is equipped with an 
arbitrary order. In this way, we can access the $n^{th}$ member of a set $s$ by 
writing $s_n$. In Figure~\ref{fig:choice}, $(G(a_j))_i$ is written a $g_j^i$.

@@

Starting from the bottom parent node, we can construct children of a goal of 
this node by choosing a goal $g_i^0$ on this node and choosing tactics 
$t_1,t_2,\ldots$ to 
apply to this goal. Here the result produces new nodes each containing a list 
of goals to be proven called obligations. In each node of the proof tree, 
a goal can either be 
solved, open or pending (see Definition~\ref{def:solved}).

In the following, concepts and properties are defined in the context of a 
search tree $\mathfrak{T}$.

@@

\section{Prediction}\label{s:prediction}
The learning-based selection of relevant lemmas significantly improves the 
automation for hammers~\cite{BlanchetteGKKU16}. In \tactictoe, we adapt 
one of the strongest lemma selection 
methods: the modified distance-weighted \emph{$k$ nearest-neighbour} 
classifier~\cite{ckju-pxtp13,DudaniS76} to predict three different objects: 
 tactics, theorems and list of goals. The number of predicted objects $k$ is 
 called the prediction radius.

@@

We now concentrate on the definition a reasonable evaluation function for the 
output of tactics. Theoretically, an interesting list of goal is one 
that has a short proof, and we could measure the likelihood of being 
provable in a number of steps from previous proof search. However, this 
approach produces a lot of negatives\todo{CK: cite millions lemma mining 
or/and deep math}. So to reduce this number, we limit 
ourselves to collect output of tactics tested during orthogonalization. We 
declare a list of goals $l$ to be positive if it is has been produced by the 
winner of an orthogonalization competition. Our predictor with radius $r$ 
produces a set of $r$ list of goals 
closest to $l$. The 
number of positive example in this set is noted $Positive(r,l)$.
During the proof search, we evaluate nodes through the list of goals they 
contain using the following evaluation function:
\begin{align*}
PriorEvaluation_r (a_i) &= \frac{Positive(r,G(a_i))}{r}\\
\end{align*}

@@

Such a prover already exists in \holfour and is called \metis. It is already 
recognized by the tactic selection mechanism in \tactictoe and can have its 
premises predicted dynamically when in an abstracted form. Nevertheless,
we think that the performance of \tactictoe can be boosted by giving the ATP 
\metis a special status. This arrangement consists in always predicting 
premises for \metis, giving it a slightly higher timeout and 
always trying it first on each node. These modifications only induce an
overhead linear on the number nodes as a general purpose prover does not create
goal obligations.
